---
title: "VA Generator Fleet Comparison: EI vs EIA"
author: "Sara Abril-Guevara"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    toc: true
    toc_float: true
    df_print: paged
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(knitr)
library(ggplot2)
```

## Introduction

This document compares generator fleets between the Energy Information Administration (EIA) and Energy Exemplar (EI) datasets. The comparison is based on geographic coordinates (latitude and longitude) with a configurable tolerance.


## Define Comparison Function





```{r}
library(dplyr)
library(readr)


#' Group generators by geographic location and aggregate their capacities
#'
#' @param gens data.frame: Generator dataframe
#' @param lat_col character: Name of latitude column
#' @param lon_col character: Name of longitude column
#' @param capacity_col character: Name of capacity column
#' @param name_col character: Name of generator name column
#'
#' @return data.frame: Grouped generators with aggregated capacity and generator lists
group_generators_by_location = function(gens, lat_col, lon_col, capacity_col, name_col) {
  grouped = gens %>%
    group_by(across(all_of(c(lat_col, lon_col)))) %>%
    summarise(
      gen_names = paste(get(name_col), collapse = "; "),
      n_generators = n(),
      total_capacity = sum(get(capacity_col), na.rm = TRUE),
      first_gen_name = first(get(name_col)),
      .groups = 'drop'
    )
  
  return(grouped)
}

#' Compare generators between EI and EIA datasets using geographic coordinates
#' First groups generators by location, then matches locations
#'
#' @param ei_gens data.frame: EI generators dataframe (must have Lat, Lon, Rating columns)
#' @param eia_gens data.frame: EIA generators dataframe (must have lat, lon, nameplate_capacity columns)
#' @param coord_tolerance numeric: Maximum distance in degrees for matching (default: 0.02 ≈ 2.2km)
#'
#' @return list: Contains grouped and detailed views of matches, plus summary statistics
compare_generator_fleets = function(ei_gens, eia_gens, coord_tolerance = 0.2) {
  
  # Step 1: Group generators by location
  cat("\nGrouping generators by location...\n")
  
  # Group EI generators
  ei_grouped = ei_gens %>%
    group_by(Lat, Lon) %>%
    summarise(
        GenNames         = paste(GenName, collapse = "; "),
        BusNumbers       = paste(BusNumber, collapse = "; "),
        BusNames         = paste(BusName, collapse = "; "),
        n_ei_generators  = n(),
        Rating           = sum(Rating),
        Available        = all(Available),
        Type             = first(Type),
        .groups          = 'drop'
    )
  
  cat("  EI:", nrow(ei_gens), "generators →", nrow(ei_grouped), "unique locations\n")
  
  # Group EIA generators
  eia_grouped = eia_gens %>%
    group_by(lat, lon) %>%
    summarise(
      plant_names = paste(plant, collapse = "; "),
      generator_ids = paste(`generator ID`, collapse = "; "),
      n_eia_generators = n(),
      nameplate_capacity = sum(nameplate_capacity),
      technology = first(technology),
      ba = first(ba),
      .groups = 'drop'
    )
  
  cat("  EIA:", nrow(eia_gens), "generators →", nrow(eia_grouped), "unique locations\n")
  
  # Step 2: Match locations
  cat("\nMatching locations (tolerance:", coord_tolerance, "°)...\n")
  
  # Function to find closest match
  find_closest_match = function(lat, lon, df, tolerance) {
    distances = sqrt((df$lat - lat)^2 + (df$lon - lon)^2)
    min_dist = min(distances)
    idx = which.min(distances)
    
    if (min_dist <= tolerance) {
      return(idx)
    } else {
      return(NA)
    }
  }
  
  # Match EI locations to EIA locations
  ei_grouped$eia_match_idx = sapply(1:nrow(ei_grouped), function(i) {
    find_closest_match(ei_grouped$Lat[i], ei_grouped$Lon[i], 
                      eia_grouped, coord_tolerance)
  })
  
  # Step 3: Create matched dataset
  in_both_grouped = ei_grouped %>% 
    filter(!is.na(eia_match_idx))
  
  if (nrow(in_both_grouped) > 0) {
    # Add EIA data
    in_both_grouped$eia_lat = eia_grouped$lat[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_lon = eia_grouped$lon[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_plant_names = eia_grouped$plant_names[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_generator_ids = eia_grouped$generator_ids[in_both_grouped$eia_match_idx]
    in_both_grouped$n_eia_generators = eia_grouped$n_eia_generators[in_both_grouped$eia_match_idx]
    in_both_grouped$eia_capacity = eia_grouped$nameplate_capacity[in_both_grouped$eia_match_idx]
    
    # Calculate differences
    in_both_grouped$capacity_diff = in_both_grouped$Rating - in_both_grouped$eia_capacity
    in_both_grouped$capacity_abs_diff = abs(in_both_grouped$capacity_diff)
    
    # Percentage difference with respect to EIA capacity
    in_both_grouped$capacity_pct_diff = (in_both_grouped$Rating - in_both_grouped$eia_capacity) / 
                                         in_both_grouped$eia_capacity * 100
    in_both_grouped$capacity_abs_pct_diff = abs(in_both_grouped$capacity_pct_diff)
    in_both_grouped$capacity_ratio = in_both_grouped$Rating / in_both_grouped$eia_capacity
    
    
    
    # Rename for clarity
    in_both_grouped = in_both_grouped %>%
      rename(ei_lat = Lat, ei_lon = Lon)
  }
  
  # Step 4: Create detailed view (expand to individual generators)
  in_both_detailed = data.frame()
  
  if (nrow(in_both_grouped) > 0) {
    for (i in 1:nrow(in_both_grouped)) {
      row = in_both_grouped[i, ]
      
      # Get individual EI generators at this location
      ei_at_location = ei_gens %>%
        filter(Lat == row$ei_lat & Lon == row$ei_lon)
      
      # Add aggregated matching info to each
      ei_at_location$eia_plant_names = row$eia_plant_names
      ei_at_location$eia_generator_ids = row$eia_generator_ids
      ei_at_location$n_eia_generators = row$n_eia_generators
      ei_at_location$n_ei_at_location = row$n_ei_generators
      ei_at_location$eia_capacity = row$eia_capacity
      ei_at_location$total_ei_capacity_at_location = row$Rating
      ei_at_location$eia_lat = row$eia_lat
      ei_at_location$eia_lon = row$eia_lon
      ei_at_location$capacity_diff = row$capacity_diff
      ei_at_location$capacity_abs_diff = row$capacity_abs_diff
      
      # Rename lat/lon
      ei_at_location = ei_at_location %>%
        rename(ei_lat = Lat, ei_lon = Lon)
      
      in_both_detailed = bind_rows(in_both_detailed, ei_at_location)
    }
  }
  
  # Step 5: Identify unmatched generators
  in_ei_only_grouped = ei_grouped %>% 
    filter(is.na(eia_match_idx))
  
  matched_eia_indices = in_both_grouped$eia_match_idx[!is.na(in_both_grouped$eia_match_idx)]
  in_eia_only_grouped = eia_grouped %>% 
    filter(!(row_number() %in% matched_eia_indices))
  
  # Calculate summary statistics
  total_ei_capacity = sum(ei_gens$Rating, na.rm = TRUE)
  total_eia_capacity = sum(eia_gens$nameplate_capacity, na.rm = TRUE)
  matched_ei_capacity = if(nrow(in_both_grouped) > 0) sum(in_both_grouped$Rating, na.rm = TRUE) else 0
  matched_eia_capacity = if(nrow(in_both_grouped) > 0) sum(in_both_grouped$eia_capacity, na.rm = TRUE) else 0
  mean_capacity_diff = if(nrow(in_both_grouped) > 0) mean(in_both_grouped$capacity_diff, na.rm = TRUE) else 0
  mean_capacity_abs_diff = if(nrow(in_both_grouped) > 0) mean(in_both_grouped$capacity_abs_diff, na.rm = TRUE) else 0
  
  # Print summary
  cat("\n", paste(rep("=", 60), collapse = ""), "\n")
  cat("Fleet Comparison Results\n")
  cat(paste(rep("=", 60), collapse = ""), "\n")
  cat("Matched locations:", nrow(in_both_grouped), "\n")
  if(nrow(in_both_grouped) > 0) {
    cat("  Total EI generators at matched locations:", sum(in_both_grouped$n_ei_generators), "\n")
    cat("  Total EIA generators at matched locations:", sum(in_both_grouped$n_eia_generators), "\n")
  }
  cat("Unmatched EI locations:", nrow(in_ei_only_grouped), "\n")
  cat("Unmatched EIA locations:", nrow(in_eia_only_grouped), "\n")
  cat("\nCapacity Summary:\n")
  cat("  Total EI capacity:", round(total_ei_capacity, 2), "MW\n")
  cat("  Total EIA capacity:", round(total_eia_capacity, 2), "MW\n")
  cat("  Matched EI capacity:", round(matched_ei_capacity, 2), "MW\n")
  cat("  Matched EIA capacity:", round(matched_eia_capacity, 2), "MW\n")
  cat("  Mean capacity difference:", round(mean_capacity_diff, 2), "MW\n")
  cat("  Mean absolute capacity difference:", round(mean_capacity_abs_diff, 2), "MW\n")
  cat(paste(rep("=", 60), collapse = ""), "\n\n")
  
  # Return results as a list
  return(list(
    in_both_grouped = in_both_grouped,
    in_both_detailed = in_both_detailed,
    in_ei_only_grouped = in_ei_only_grouped,
    in_eia_only_grouped = in_eia_only_grouped,
    ei_grouped = ei_grouped,
    eia_grouped = eia_grouped,
    summary = list(
      total_ei_capacity = total_ei_capacity,
      total_eia_capacity = total_eia_capacity,
      matched_ei_capacity = matched_ei_capacity,
      matched_eia_capacity = matched_eia_capacity,
      mean_capacity_diff = mean_capacity_diff,
      mean_capacity_abs_diff = mean_capacity_abs_diff,
      n_matched_locations = nrow(in_both_grouped),
      n_ei_only_locations = nrow(in_ei_only_grouped),
      n_eia_only_locations = nrow(in_eia_only_grouped)
    )
  ))
}


```

# Read all the gens

```{r}
VA_ei_gens = read_csv("~/Downloads/virginia_resources.csv")
eia_rd = read_csv("renewable_VA_EIA.csv")
```

### Exploring the datasets

```{r}
unique(VA_ei_gens$generator_type)
unique(eia_rd$technology)
```


```{r}
#' Compare generator fleets for a specific technology type
#'
#' @param ei_gens data.frame: Full EI generators dataframe (virginia_resources.csv)
#' @param eia_gens data.frame: Full EIA generators dataframe
#' @param type character: "Solar", "Wind", "Hydro", "Thermal", "Storage"
#' @param coord_tolerance numeric: tolerance in degrees (default 0.2)
compare_by_type = function(ei_gens, eia_gens, type = "Solar", coord_tolerance = 0.2) {
  
  # ── 1. Type → filter mappings ──────────────────────────────────────────────
  ei_type_map = list(
    "Solar"   = c("Solar"),
    "Wind"    = c("Wind"),
    "Hydro"   = c("Hydro"),
    "Thermal" = c("ThermalStandard", "ThermalMultiStart"),
    "Storage" = c("Storage")
  )
  
  eia_type_map = list(
    "Solar"   = c("Solar Photovoltaic"),
    "Wind"    = c("Onshore Wind Turbine", "Offshore Wind Turbine"),
    "Hydro"   = c("Conventional Hydroelectric", "Run-of-River Hydroelectric"),
    "Thermal" = c("Natural Gas Steam Turbine", "Natural Gas Fired Combined Cycle",
                  "Natural Gas Fired Combustion Turbine", "Coal Steam Turbine"),
    "Storage" = c("Batteries")
  )
  
  if (!type %in% names(ei_type_map)) {
    stop("Invalid type. Choose from: ", paste(names(ei_type_map), collapse = ", "))
  }
  
  # ── 2. Filter by type ──────────────────────────────────────────────────────
  ei_filtered = ei_gens %>% filter(generator_type %in% ei_type_map[[type]])
  eia_filtered = eia_gens %>% filter(technology %in% eia_type_map[[type]])
  
  cat("=== Type:", type, "===\n")
  cat("EI generators:  ", nrow(ei_filtered), "\n")
  cat("EIA generators: ", nrow(eia_filtered), "\n")
  
  if (nrow(ei_filtered) == 0 || nrow(eia_filtered) == 0) {
    warning("No generators found for type: ", type)
    return(NULL)
  }
  
  # ── 3. Rename EI columns to match compare_generator_fleets expectations ────
  ei_renamed = ei_filtered %>%
    rename(
      GenName   = name,
      BusNumber = bus_number,
      BusName   = bus,
      Lat       = lat,
      Lon       = lon,
      Rating    = rating,
      Available = available,
      Type      = generator_type
      # Rating stays as Rating ✅
    )
  
  # ── 4. Run comparison ──────────────────────────────────────────────────────
  result = compare_generator_fleets(
    ei_gens        = ei_renamed,
    eia_gens       = eia_filtered,
    coord_tolerance = coord_tolerance
  )
  
  return(result)
}
```

### Solar

```{r}
# Solar comparisons
solar_results = compare_by_type(VA_ei_gens, eia_rd, type = "Solar", coord_tolerance = 0.2)

# Wind comparison
#wind_results  = compare_by_type(VA_ei_gens, eia_rd, type = "Wind",  coord_tolerance = 0.2)

# Hydro comparison
#hydro_results = compare_by_type(VA_ei_gens, eia_rd, type = "Hydro", coord_tolerance = 0.2)

# Access results
solar_results$in_both_grouped%>%
  dplyr::select(n_eia_generators, eia_plant_names, n_ei_generators, BusNames,
                Rating, eia_capacity, 
                capacity_diff, capacity_abs_diff,
                capacity_pct_diff, capacity_abs_pct_diff,   
                ei_lat, eia_lat, ei_lon, eia_lon) %>%
  dplyr::arrange(desc(capacity_abs_pct_diff))  # sort by % diff    # matched locations
#solar_results$in_ei_only_grouped  # only in EI
#solar_results$in_eia_only_grouped # only in EIA
```

### Wind

```{r}
# Wind comparison
wind_results  = compare_by_type(VA_ei_gens, eia_rd, type = "Wind",  coord_tolerance = 0.2)

wind_results$in_both_grouped%>%
  dplyr::select(n_eia_generators, eia_plant_names, n_ei_generators, BusNames,
                Rating, eia_capacity, 
                capacity_diff, capacity_abs_diff,
                capacity_pct_diff, capacity_abs_pct_diff,   
                ei_lat, eia_lat, ei_lon, eia_lon) %>%
  dplyr::arrange(desc(capacity_abs_pct_diff))  # sort by % diff    # matched locations

ei_wind_gens = VA_ei_gens %>% filter(Type == "Wind")
eia_wind_gens = eia_rd %>% filter(technology == "Offshore Wind Turbine")



```

```{r}
# Hydro comparison
eia_hydro = read_csv("~/Documents/GitHub/va_updates/va_hydro_EIA.csv")
hydro_results = compare_by_type(VA_ei_gens, eia_hydro, type = "Hydro", coord_tolerance = 0.2)

```



## Solar

```{r}
ei_solar_gens = VA_ei_gens %>% filter(Type == "Solar")
eia_solar_gens = eia_rd %>% filter(technology == "Solar Photovoltaic")

str(ei_solar_gens)
str(eia_solar_gens)
```



```{r}


renewable_results = compare_generator_fleets(ei_rd_gens, eia_rd, coord_tolerance = 0.1)

# View grouped results (one row per location)
renewable_results$in_both_grouped %>%
  dplyr::select(n_eia_generators, eia_plant_names, n_ei_generators, BusNames,
                Rating, eia_capacity, 
                capacity_diff, capacity_abs_diff,
                capacity_pct_diff, capacity_abs_pct_diff,   
                ei_lat, eia_lat, ei_lon, eia_lon) %>%
  dplyr::arrange(desc(capacity_abs_pct_diff))  # sort by % diff 
  

# View detailed results (one row per EI generator)
renewable_results$in_both_detailed %>%
  dplyr::select(GenName, BusNumber, n_ei_at_location, eia_plant_names, n_eia_generators,
         Rating, eia_capacity, capacity_abs_diff) %>%
  dplyr::arrange(desc(capacity_abs_diff))
```


```{r}
library(ggplot2)

ggplot(renewable_results$in_both_grouped,
       aes(x = eia_capacity, y = Rating)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  scale_x_log10() +
  scale_y_log10()
```




### Thermal

```{r}

library(readxl)

thermal_eia = read_excel("Thermal_EIA.xlsx")
ei_thermal = VA_ei_gens %>%
  dplyr::filter(Type %in% c("ThermalStandard", "ThermalMultiStart"))
ei_thermal = ei_thermal %>% filter(!is.na(Lat) & !is.na(Lon))
thermal_eia = thermal_eia %>% filter(!is.na(lat) & !is.na(lon))

```


```{r}
thermal_results = compare_generator_fleets(
  ei_thermal,
  thermal_eia,
  coord_tolerance = 0.1
)
```

```{r}
# View grouped results (one row per location)
thermal_results$in_both_grouped %>%
  dplyr::select(n_eia_generators, eia_plant_names, n_ei_generators, BusNames,
                Rating, eia_capacity, 
                capacity_diff, capacity_abs_diff,
                capacity_pct_diff, capacity_abs_pct_diff,   
                ei_lat, eia_lat, ei_lon, eia_lon) %>%
  dplyr::arrange(desc(capacity_abs_pct_diff))  # sort by % diff 
  

# View detailed results (one row per EI generator)
thermal_results$in_both_detailed %>%
  dplyr::select(GenName, BusNumber, n_ei_at_location, eia_plant_names, n_eia_generators,
         Rating, eia_capacity, capacity_abs_diff) %>%
  dplyr::arrange(desc(capacity_abs_diff))
```



# Hydro

```{r}

library(readxl)

hydro_eia = read_csv("va_hydro_EIA.csv")
ei_thermal = VA_ei_gens %>%
  dplyr::filter(Type %in% c("ThermalStandard", "ThermalMultiStart"))
ei_thermal = ei_thermal %>% filter(!is.na(Lat) & !is.na(Lon))
thermal_eia = thermal_eia %>% filter(!is.na(lat) & !is.na(lon))

```


```{r}
thermal_results = compare_generator_fleets(
  ei_thermal,
  thermal_eia,
  coord_tolerance = 0.1
)
```



## Load Data

```{r load-data, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE, results='hide'}
# Load EI generators
VA_ei_gens = read_csv("VA_original_EI.csv")

# Filter renewable dispatch generators
ei_rd_gens = VA_ei_gens %>% 
  filter(Type == "RenewableDispatch")

# Load EIA renewable generators
eia_rd = read_csv("renewable_VA_EIA.csv")
```

## Data Overview

### EI Renewable Generators

```{r ei-overview}
cat("Total EI renewable generators:", nrow(ei_rd_gens), "\n")
cat("Total EI renewable capacity:", round(sum(ei_rd_gens$Rating), 2), "MW\n")
cat("Available generators:", sum(ei_rd_gens$Available), "\n")
cat("Unavailable generators:", sum(!ei_rd_gens$Available), "\n")
```

**Total EI renewable generators:** `r nrow(ei_rd_gens)`

**Total EI renewable capacity:** `r round(sum(ei_rd_gens$Rating), 2)` MW

**Available generators:** `r sum(ei_rd_gens$Available)`

**Unavailable generators:** `r sum(!ei_rd_gens$Available)`

### EIA Renewable Generators

```{r eia-overview}
cat("Total EIA renewable generators:", nrow(eia_rd), "\n")
cat("Total EIA renewable capacity:", round(sum(eia_rd$nameplate_capacity), 2), "MW\n")
```

## Run Comparison

We use a coordinate tolerance of 0.02 degrees (approximately 2.2 km) to match generators.

```{r run-comparison}
renewable_results = compare_generator_fleets(
  ei_rd_gens, 
  eia_rd, 
  coord_tolerance = 0.02
)
```

## Results

### Summary Statistics

```{r summary-table}
summary_df = data.frame(
  Metric = c("Generators in both datasets", 
             "Generators in EI only", 
             "Generators in EIA only",
             "Total EI capacity (MW)",
             "Total EIA capacity (MW)",
             "Matched EI capacity (MW)",
             "Matched EIA capacity (MW)",
             "Mean capacity absolute difference (MW)"),
  Value = c(
    renewable_results$summary$n_matched,
    renewable_results$summary$n_ei_only,
    renewable_results$summary$n_eia_only,
    round(renewable_results$summary$total_ei_capacity, 2),
    round(renewable_results$summary$total_eia_capacity, 2),
    round(renewable_results$summary$matched_ei_capacity, 2),
    round(renewable_results$summary$matched_eia_capacity, 2),
    round(renewable_results$summary$mean_capacity_abs_diff, 2)
  )
)

kable(summary_df, caption = "Renewable Generator Fleet Comparison Summary")
```

### Matched Generators

```{r matched-table}
if (nrow(renewable_results$in_both) > 0) {
  matched_display = renewable_results$in_both %>%
    dplyr::select(BusNumber, GenName, BusName, eia_plant, eia_n_generators,
                  Rating, eia_capacity, capacity_diff, capacity_abs_diff,
                  ei_lat, ei_lon, eia_lat, eia_lon, Available) %>%
    arrange(desc(capacity_abs_diff))
  
  kable(matched_display, 
        caption = "Sample of Matched Generators (first 20)",
        digits = 4)
}

```

```{r}
# Run comparison
renewable_results = compare_generator_fleets(ei_rd_gens, eia_rd, coord_tolerance = 0.02)

# View grouped results (one row per location)
renewable_results$in_both_grouped %>%
  select(ei_lat, ei_lon, n_ei_generators, eia_plant, eia_n_generators, 
         Rating, eia_capacity, capacity_diff, capacity_abs_diff) %>% arrange(desc(capacity_abs_diff)) 

# View detailed results (one row per EI generator)
renewable_results$in_both_detailed %>%
  dplyr::select(GenName, BusNumber, n_ei_at_location, eia_plant, eia_n_generators,
         Rating, eia_capacity, capacity_diff,capacity_abs_diff, ei_lat, ei_lon) %>%
  dplyr::arrange(desc(capacity_abs_diff))
```



### Capacity Difference Distribution

```{r capacity-plot, fig.width=10, fig.height=6}
if (nrow(renewable_results$in_both) > 0) {
  ggplot(renewable_results$in_both, aes(x = capacity_abs_diff)) +
    geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
    labs(title = "Distribution of Capacity Absolute Differences (EI - EIA)",
         x = "Capacity Difference (MW)",
         y = "Count") +
    theme_minimal()
}
```

### Geographic Distribution

```{r geo-plot, fig.width=12, fig.height=8}
if (nrow(renewable_results$in_both) > 0) {
  ggplot() +
    geom_point(data = renewable_results$in_both, 
               aes(x = ei_lon, y = ei_lat), 
               color = "blue", alpha = 0.6, size = 3) +
    geom_point(data = renewable_results$in_both, 
               aes(x = eia_lon, y = eia_lat), 
               color = "red", alpha = 0.6, size = 3) +
    geom_segment(data = renewable_results$in_both,
                 aes(x = ei_lon, y = ei_lat, 
                     xend = eia_lon, yend = eia_lat),
                 alpha = 0.3) +
    labs(title = "Geographic Matching of Generators",
         subtitle = "Blue = EI location, Red = EIA location",
         x = "Longitude",
         y = "Latitude") +
    theme_minimal()
}
```

### EI-Only Generators

These generators exist in EI but not in EIA (potentially unavailable or retired):

```{r ei-only}
cat("Number of EI-only generators:", nrow(renewable_results$in_ei_only), "\n")
cat("Unavailable generators in EI-only:", 
    sum(!renewable_results$in_ei_only$Available), "\n")

if (nrow(renewable_results$in_ei_only) > 0) {
  ei_only_display = renewable_results$in_ei_only %>%
    dplyr::select(GenName, Rating, Lat, Lon, Available) %>%
    head(20)
  
  kable(ei_only_display, 
        caption = "Sample of EI-Only Generators (first 20)")
}
```

### EIA-Only Generators

These generators exist in EIA but not in EI:

```{r eia-only}
cat("Number of EIA-only generators:", nrow(renewable_results$in_eia_only), "\n")

if (nrow(renewable_results$in_eia_only) > 0) {
  eia_only_display = renewable_results$in_eia_only %>%
    dplyr::select(plant, technology, nameplate_capacity, lat, lon) %>%
    head(20)
  
  kable(eia_only_display, 
        caption = "Sample of EIA-Only Generators (first 20)")
}
```

## Export Results

```{r export-results, eval=FALSE}
# # Export matched generators
# write_csv(renewable_results$in_both, "matched_renewable_generators.csv")
# 
# # Export EI-only generators
# write_csv(renewable_results$in_ei_only, "ei_only_renewable_generators.csv")
# 
# # Export EIA-only generators
# write_csv(renewable_results$in_eia_only, "eia_only_renewable_generators.csv")
```

## Conclusion

This comparison identified:

- **`r renewable_results$summary$n_matched`** matched generators between EI and EIA
- **`r renewable_results$summary$n_ei_only`** generators only in EI (may be unavailable)
- **`r renewable_results$summary$n_eia_only`** generators only in EIA (not in the system model)

The Mean capacity absolute difference for matched generators is **`r round(renewable_results$summary$mean_capacity_abs_diff, 2)` MW**.